---
title: "Technical Assessment: Data Analysis Exercise for Candidate Evaluation"
author: "Sebastián Sánchez Cuartas"
date: ""
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
      position: "left"
    theme: flatly
    highlight: tango
---

```{=html}
<style>
/* Mantiene la misma fuente para los blockquotes */
blockquote {
    font-family: inherit; /* Usa la misma fuente que el resto del texto */
    font-size: inherit; /* Usa el mismo tamaño de fuente que el resto del texto */
}

/* Ajusta el ancho del menú de la tabla de contenido flotante */
.tocify {
    width: 1300px; /* Ajusta este valor según sea necesario */
}
.tocify-wrapper {
    width: 1300px; /* Ajusta este valor según sea necesario */
}
</style>
```
\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

## **Assignment Overview**

The goal of this assessment is to evaluate the candidate's technical and analytical skills. The task involves analyzing a sample dataset from a client in the Consumer-Packaged Goods (CPG) sector to generate actionable insights. Here are the details provided for the assignment:

**Task Details**

-   **Objective**: Identify patterns in the data that can inform the client's marketing and supply chain strategies.
-   **Tools**: You may use R, Python, or any object-oriented programming language.
-   **Time Limit**: 3 hours. While time management is part of the evaluation, the quality of the analysis is prioritized.

**Deliverables**

-   A Jupyter Notebook or R script with the complete analysis and code.
-   A brief report (1-2 pages) summarizing findings and recommendations for the CPG client.

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

## **Technical Assessment**

**Data Exploration & Quality Control**

-   <div>

    > -   Load the dataset and explore its structure.
    >
    > -   Identify any missing values, duplicates, or inconsistencies.
    >
    > -   Provide a summary of the dataset, including variable types, record counts, and notable patterns.
    >
    > -   Ensure data quality for client presentation.

    </div>

**Data Cleaning**

-   <div>

    > -   Address missing or inconsistent values, and document the approach.
    >
    > -   Standardize names, regions, and other key fields for accurate reporting.

    </div>

**Exploratory Data Analysis (EDA)**

-   <div>

    > -   Conduct exploratory analysis to identify trends or patterns.
    >
    > -   Focus on key fields and present insights in the final report.

    </div>

**SQL Query**

-   <div>

    > Write a SQL query to extract counts by the main category.

    </div>

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

## **(1) Data Cleaning and Preparation**

```{r, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
library(dplyr)

# Load the CSV file
data <- read.csv("C:/Users/Usuario/Downloads/Take_home_data.csv", stringsAsFactors = FALSE, sep = ";")

# Summary of the data frame structure
str(data)

# Clean special characters
data <- data %>%
  mutate(across(-c(1, 2), ~ iconv(., to = "UTF-8", sub = "byte")))  

# Convert to lowercase and replace whitespace with underscores, except for columns 1 and 2, which may contain uppercase codes
data <- data %>%
  mutate(across(-c(1, 2), ~ gsub(" ", "_", tolower(.))))

# Replace commas with periods and convert to numeric
data$popularity_score <- as.numeric(gsub(",", ".", data$popularity_score))
data$sentiment_score <- as.numeric(gsub(",", ".", data$sentiment_score))
data$latitude <- as.numeric(gsub(",", ".", data$latitude))
data$longitude <- as.numeric(gsub(",", ".", data$longitude))
data$open_closed_status_confidence_score <- as.numeric(gsub(",", ".", data$open_closed_status_confidence_score))

# Replace empty strings with NA
data[data == ""] <- NA

# Convert the 'opened_on' column to date format
data$opened_on <- as.Date(data$opened_on, format = "%d/%m/%Y")

# Remove the "<a0>" character to make the hours in the corresponding columns more readable
data <- data %>%
  mutate(across(29:42, 
                ~ {gsub("<a0>", "", .)
                }))

# Check unique values in key columns to detect errors
unique(data$main_category)
unique(data$sub_category)
unique(data$neighborhood)
unique(data$city)
unique(data$state)
unique(data$country_code)
unique(data$country)
unique(data$dataplor_status)
unique(data$open_closed_status)

# Identify duplicate rows
duplicates <- data[duplicated(data), ]

# Remove duplicates
data <- data[!duplicated(data), ]

```

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

## **(2) Missing Values Visualization**

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=6, out.width="100%"}
# Visualize missing values
library(naniar)

gg_miss_var(data)

# Check the number of NA values by column
colSums(is.na(data))

```

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

## **(3) Imputation of Missing Values and Comparative Analysis**

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=6, out.width="100%"}

library(tidyr)
library(dplyr)
library(ggplot2)

#Given the number of missing values in popularity_score and sentiment_score, which is approximately one-third of the data, we perform a comparative analysis between imputed vs. non-imputed elements to make a decision.

# Impute missing values for popularity and sentiment scores directly in the dataset
data <- data %>%
  group_by(main_category) %>%
  mutate(
    popularity_score_imputed = ifelse(is.na(popularity_score), median(popularity_score, na.rm = TRUE), popularity_score),
    sentiment_score_imputed = ifelse(is.na(sentiment_score), median(sentiment_score, na.rm = TRUE), sentiment_score)
  ) %>%
  ungroup()

# Reshape data for visualization
data_comparison <- data %>%
  pivot_longer(
    cols = c(popularity_score, popularity_score_imputed, sentiment_score, sentiment_score_imputed),
    names_to = c(".value", "Type"),
    names_pattern = "(.+)_score(_imputed)?"
  )

# Create boxplots for comparison
ggplot(data_comparison, aes(x = main_category, y = popularity, fill = Type)) +
  geom_boxplot() +
  theme_minimal() +
  xlab("main_category") +
  ylab("popularity") +
  ggtitle("Distribution of popularity (Original vs Imputed)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(data_comparison, aes(x = main_category, y = sentiment, fill = Type)) +
  geom_boxplot() +
  theme_minimal() +
  xlab("main_category") +
  ylab("sentiment") +
  ggtitle("Distribution of sentiment (Original vs Imputed)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


# From the analysis above, it is concluded that the imputed data does not differ significantly from the non-imputed data. Therefore, we will proceed with the analyses using the imputed data, as it provides a larger sample size for subsequent analyses. However, it is important to note that this approach sacrifices some of the original data's variability.


```

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

## **(4) Distribution of Main Business Categories**

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=6, out.width="100%"}

# View the distribution of main business categories

ggplot(data, aes(x = main_category)) +
  geom_bar() +
  theme_minimal() +
  xlab("Main Category") +
  ylab("Frequency") +
  ggtitle("Distribution of Main Business Categories") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Justification: This plot shows the distribution of different types of businesses by main category, making it easier to identify which types of businesses are most common in the database.

# Importance for the Client: Understanding the most prevalent business categories can help the client identify segments with a high market presence and explore opportunities for expansion or improvement in less represented categories.

```

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

## **(5) Summary of Popularity and Sentiment by Business Category**

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=6, out.width="100%"}

# Select the 20 most common categories for the business_category variable
top_categories <- names(sort(table(data$business_category), decreasing = TRUE)[1:20])
data_top <- data %>%
  filter(business_category %in% top_categories)

# Summary of popularity and sentiment by business_category, including main_category
summary_popularity_sentiment <- data_top %>%
  group_by(business_category) %>%
  summarise(
    avg_popularity = mean(popularity_score, na.rm = TRUE),
    avg_sentiment = mean(sentiment_score, na.rm = TRUE),
    count = n()
  ) %>%
  left_join(data_top %>% select(business_category, main_category) %>% distinct(), by = "business_category") %>%
  arrange(desc(avg_popularity))

# View the results
print(summary_popularity_sentiment)

# Importance for the Client: This analysis allows the client to identify the most prominent categories in terms of popularity and positive perception, which is crucial for determining which areas to enhance or improve.

```

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

## **(6) Relationship Between Popularity and Sentiment for Top Business Categories**

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=6, out.width="100%"}

# Faceted plot for the main business_categories using jitter to avoid overlap
ggplot(data_top, aes(x = popularity_score, y = sentiment_score)) +
  geom_jitter(alpha = 0.5, width = 0.02, height = 0.02) +  # Adjust 'width' and 'height' as needed
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  facet_wrap(~ business_category) +
  theme_minimal() +
  xlab("Popularity Score") +
  ylab("Sentiment Score") +
  ggtitle("Relationship between Popularity and Sentiment (Top 20 Categories) with Jitter")

# Justification: Filtering the 20 most common categories helps focus the analysis on those with greater relevance. This avoids visual overload and facilitates the interpretation of the graphs and results.

# Importance for the Client: This helps identify specific patterns between popularity and sentiment for each business category, showing whether there are categories with high popularity and low sentiment (or vice versa), which could be useful for adjusting marketing and customer service strategies.

```

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

## **(7) Distribution of Popularity and Sentiment Scores by Business Category**

```{r, warning=FALSE, message=FALSE, fig.width=9, fig.height=6, out.width="100%"}

# Complementary boxplot to visualize the distribution of popularity and sentiment scores by business category
ggplot(data_top, aes(x = business_category, y = popularity_score)) +
  geom_boxplot() +
  theme_minimal() +
  coord_flip() +  # Rotate the boxplot for better visualization
  xlab("Business Category") +
  ylab("Popularity Score") +
  ggtitle("Distribution of Popularity Scores by Business Category")

ggplot(data_top, aes(x = business_category, y = sentiment_score)) +
  geom_boxplot() +
  theme_minimal() +
  coord_flip() +  # Rotate the boxplot for better visualization
  xlab("Business Category") +
  ylab("Sentiment Score") +
  ggtitle("Distribution of Sentiment Scores by Business Category")

```

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

## **(8) Analysis of Popularity and Sentiment Based on Business Open/Closed Status**

```{r, warning=FALSE, message=FALSE}

# Popularity and sentiment based on open or closed status, adjusted to two levels "open" and "closed" to perform a t-test that allows understanding the significance of scores regarding the likelihood of the business being open or closed

data_filtered <- data %>%
  filter(!is.na(open_closed_status)) %>%
  mutate(open_closed_status = ifelse(open_closed_status %in% c("permanently_closed", "temporarily_closed"), "closed", open_closed_status))

# T-tests for popularity and sentiment by open/closed status
t_test_popularity <- t.test(popularity_score ~ open_closed_status, data = data_filtered)
print(t_test_popularity)

t_test_sentiment <- t.test(sentiment_score ~ open_closed_status, data = data_filtered)
print(t_test_sentiment)



```

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

## **(9) SQL Query**

```{r, warning=FALSE, message=FALSE}

# Since the dataset is not very large, I can perform the query in R using the sqldf package.

library(sqldf)

# Ejecutar la consulta SQL usando sqldf
result <- sqldf("SELECT main_category, COUNT(*) AS category_count
                 FROM data
                 GROUP BY main_category")

# Ver el resultado
print(result)

```

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

## **(10) Data Analysis Report**

This report is structured according to the points enumerated in the R code and corresponding graphics. The following presents the results and analysis:

1.      **Data Cleaning and Preparation**\
Data cleaning and preparation were performed to ensure quality and consistency. This included removing duplicates, normalizing formats, and converting columns to appropriate data types.

2.      **Missing Values Visualization**\
A visualization of missing values in the dataset was generated. This helped identify patterns of missing data, allowing for a quick assessment of the magnitude of missing values across different variables, which aids in planning imputation strategies.

3.      **Imputation of Missing Values and Comparative Analysis**\
Imputation methods were applied to handle missing values, such as median imputation for the variables *popularity_score* and *sentiment_score*, which are crucial for understanding the success of each business. A comparative analysis was conducted before and after the imputation to assess the impact of this technique on the distribution of metrics. It was decided to proceed with the imputed data, accepting a slight sacrifice in variability.

4.      **Distribution of Main Business Categories**\
The distribution of main business categories was examined, providing insights into the variety and concentration of businesses in each category. This helps identify the most representative categories that require more attention, with *professional_services* being the most frequent. To gain deeper insights, the *business_category* variable was further analyzed, establishing a relationship with the imputed variables: *popularity_score* and *sentiment_score*.

5.      **Summary of Popularity and Sentiment by Business Category**\
A summary was created detailing popularity and sentiment by business category. This analysis reveals how these metrics are related, providing valuable insights into which categories have greater recognition and a better overall perception, as well as identifying their respective *main_category*.

6.      **Relationship Between Popularity and Sentiment for Top Business Categories**\
The relationship between popularity and sentiment was explored for the top business categories. The 20 most important categories were selected, and the faceted plot revealed that most categories show low popularity with high sentiment scores. This suggests the need for strategies to boost popularity, leveraging the positive sentiment already present.

7.      **Distribution of Popularity and Sentiment Scores by Business Category**\
The distribution of *popularity_score* and *sentiment_score* by business category was visualized. This approach provides a clearer perspective on how these metrics are distributed, visually confirming the observations from the previous analysis.

8.      **Analysis of Popularity and Sentiment Based on Business Open/Closed Status**\
An analysis of popularity and sentiment based on the open or closed status of businesses was conducted using t-tests. The results showed:

> | o    **T-Test for Popularity**: A significant p-value (\< 0.05) indicates a difference between the popularity of open businesses (0.3173) and closed businesses (0.2744), suggesting that higher popularity scores correlate with a greater likelihood of remaining open.

> | o    **T-Test for Sentiment**: A very low p-value indicates that open businesses have a higher average sentiment score (0.7637) compared to closed businesses (0.6738), implying that businesses with a positive perception are more likely to remain operational.

Based on this, it is recommended that the client focus on improving the perception of their businesses, as this could increase the likelihood of remaining open.

## **(11) Conclusion**

In summary, the analysis reveals that both popularity and positive perception are critical factors for business sustainability. Implementing strategies that encourage higher popularity and a positive perception can significantly contribute to the operational continuity of businesses in the market. Additionally, considering the geographic coordinates available in the dataset, spatial analysis could provide further insights into regional trends and opportunities.
